The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-05-11 Paper/jaccardbaetica.tex
add new changes britannia germania

The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}.

The majority of olive oil amphorae produced in this province and shipped through the provinces belong to Dressel 20 type \citep{dressel_ricerche_1878,
millet_anforas_1998}. In particular, the importance of olive oil production could be explained by the fact that we detected amphorae distributed in different Roman provinces such as \textit{Germania}, \textit{Britannia} and \textit{Gallia}. This amphora was commonly associated with the  transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.

Most Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti} with different information still unknown. Indeed, most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.

@@

Nevertheless, some challenges remain under discussion such as how this production was organised and whether it is possible to distinguish production patterns in the olive oil trade. Our questions will be focused on the distribution of amphoric stamps. Did they follow a distribution pattern? Did stamps share the same workshop? Neither written records have been found that it could explain the economic role of \textit{Baetica} province in the Roman organisation. Indeed, archaeological evidence shows a highly specialised production with a long activity in this specific area with apparently few changes \citep{remesal_anforas_2004}.

@@

\section{Case study: consumption centers}
%\xavi{este título de sección debería cambiarse por algo más descriptivo. Case studies o algo asínnn? Por otra parte, igual podrías dividirlo en 2 subsections}

@@

The conquest of new provinces allowed to the Roman Empire the arrival of new resources. This lead to a progressive change in the economic and social structure. As a result, Augusto's administration created the figure of the \textit{praefectura annonae} for the supply of wheat. The role of the \textit{praefectura annonae} was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens \citep{remesal_annona_1986,remesal_concierto}.

The consumption of olive oil in \textit{Britannia} was residual until starting the Roman conquest \citep{funari_corpus_1996,
carreras_abastecimiento_2003}.

It is well known that this product was not frequently consumed by indigenous population \citep[161]{monfort_britanniaen_1998}. The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[ 1]{carreras_britannia_1998}. Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions. This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman army.

>>>>

The chronology in the workshops is widely diverse from the first to the third centuries AD \citep{millet_anforas_1998,rodriguez_baetican_1998,chic2005comercio}. 
This fact is shown by the archaeological evidence that displays a highly specialised production with a long activity with apparently few changes \citep{remesal_anforas_2004}.

The majority of olive oil amphorae produced in this province belong to Dressel 20 type\citep{dressel_ricerche_1878,
millet_anforas_1998}. Dressel 20 amphorae were shipped and distributed from Baetica throughout western Europe.

Dressel 20 contains many information still unknown. A high percentage of  Dressel 20 were marked in stamps, inked in \textit{tituli picti} and incise in \textit{graffiti}. Indeed, most of the inscriptions on the amphora have not enough analysed due to the fragmentation of the material or the shortage of samples \citep{aguilera_evolucion_2007,rovira_guardiola_grafitos_2007}.

@@

Nevertheless, some challenges remain under discussion such as how this production was organised and whether it is possible to distinguish production patterns in the olive oil trade. Our questions will be focused on the distribution of amphoric stamps. Did they follow a distribution pattern? Did stamps share the same workshop? Neither written records have been found that it could explain the economic role of \textit{Baetica} province in the Roman organisation.

@@

\section{Case study: consumption centers}

The conquest of new provinces allowed to the Roman Empire the arrival of new resources through Mediterranean and Atlantic routes. This lead to a progressive change in the economic and social structure creating new trade networks for supplying Roman conquest. As a result, Augusto's administration created the figure of the \textit{praefectura annonae} for the supply of wheat. The role of the \textit{praefectura annonae} was mainly to provide though \textit{frumentationes} a quantity per month of wheat to Roman citizens \citep{remesal_annona_1986,remesal_concierto}.

\textit{Annona} could have also included other goods for the supply of Roman legions such as olive oil according for some hypotheses \citep{remesal_annona_1986,remesal_annona_1990}

The importance of olive oil production could be explained by the fact that we detected a massive amphora distribution in different Roman provinces such as \textit{Germania} and \textit{Britannia}. This amphora has been commonly associated with the  transportation of Baetican olive oil for supplying military camps and civil settlements \citep{berni_millet_epigrafianforica_2008}.

The concentration of amphorae in eminently militarised areas suggests a possible existence of a broad supply mechanism, being accurate to imagine a possible organisation in charge of its distribution from military camps to civil settlements\citep{remesal_annona_1986, carreras_britannia_1998}.

The military consumption of olive oil in Roman provinces could be related to two aspects: a) cultural consumption whose product is consumed by cultural reasons such as to identity or habit and b) economical reasons where olive oil distribution is based on the transport costs unlike other products \citep[69-70]{carreras_britannia_1998}.

\maria{falta conexión entre una cosa y hacer una breve introducción de lo siguiente britannia y germania}

%\xavi{este título de sección debería cambiarse por algo más descriptivo. Case studies o algo asínnn? Por otra parte, igual podrías dividirlo en 2 subsections}

@@

The consumption of olive oil in \textit{Britannia} was residual until starting the Roman conquest \citep{funari_corpus_1996,
carreras_abastecimiento_2003}. It is well known that this product was not frequently consumed by indigenous population \citep[161]{monfort_britanniaen_1998}. The absence of olive oil importation before the conquest can be reflected by the shortage of this product until the arrival of a high percentage of Dressel 20 amphorae to \textit{Britannia} \citep[1]{carreras_britannia_1998}. Additionally, the land of \textit{Britannia} showed inadequate for the olive oil production due to low environmental conditions. This issue was necessary for creating an important export apparatus from \textit{Baetica} to supply Roman army.
